Laplacian envelope

### 2022-06-03 Notes/laplace.tex
typos

\newcommand{\aav}[1]{\mathopen{}\left\llangle #1 \right\rrangle\mathclose{}}
\newcommand{\nnrm}[1]{\mathopen{}\left\lWavy #1 \right\rWavy\mathclose{}}
%================================================================================
%
% \usepackage{adjustbox}
% \newcommand{\aligntop}[1]{\adjustbox{valign=t}{#1}}
% \newcommand{\alignmid}[1]{\adjustbox{valign=c}{#1}}
% %
% \usepackage{paralist}
% \newenvironment{myenuma}{\begin{inparaenum}[(a)]}{\end{inparaenum}}
% \newenvironment{myenumi}{\begin{inparaenum}[i)]}{\end{inparaenum}}
% \newenvironment{myenumA}{\begin{inparaenum}[\bfseries A.]}{\end{inparaenum}}
% \newenvironment{myenumI}{\begin{inparaenum}[\bfseries I.]}{\end{inparaenum}}
% \newenvironment{myenumr}{\begin{inparaenum}[1.]}{\end{inparaenum}}
% %for nested lists, \ref only produces innermost item:  (hacky)
% \makeatletter % for internal macros with @
% \renewcommand\p@enumii{}
% \makeatother
%
%================================================================================
%matrices
%vec of ones
\newcommand{\onev}{\mathbf{e}}
%mat of ones
\newcommand{\onem}{\mathbf{E}}
%Markov matrix
\newcommand{\MMm}{Q}
\newcommand{\MM}{\mathbf{\MMm}}
%prob distribution
\newcommand{\prm}{p}
\newcommand{\pr}{\mathbf{\prm}}
%equilibrium distribution
\newcommand{\eqm}{\pi}
\newcommand{\eq}{\boldsymbol{\eqm}}
%equilibrium metric
\newcommand{\Eqm}{\Pi}
\newcommand{\Eq}{\boldsymbol{\Eqm}}
%first passage times
\newcommand{\fptm}{T}
\newcommand{\fpt}{\mathbf{\fptm}}
%off-diag first passage times
\newcommand{\fptbm}{\overline{\fptm}}
\newcommand{\fptb}{\overline{\fpt}}
%fundamental matrix
\newcommand{\fundm}{Z}
\newcommand{\fund}{\mathbf{\fundm}}
%other symbols
\newcommand{\Pb}{\mathbf{P}}
\newcommand{\pib}{\boldsymbol{\pi}}
\newcommand{\arowm}{\xi}
\newcommand{\arow}{\boldsymbol{\arowm}}
\newcommand{\Lbm}{\Lambda}
\newcommand{\Lb}{\boldsymbol{\Lbm}}
\newcommand{\lbm}{\lambda}
\newcommand{\lb}{\boldsymbol{\lbm}}
\newcommand{\etwm}{\eta^w}
\newcommand{\etw}{\boldsymbol{\eta}^w}
\newcommand{\thbm}{\theta}
\newcommand{\thb}{\boldsymbol{\thbm}}
\newcommand{\ombm}{\omega}
\newcommand{\omb}{\boldsymbol{\ombm}}
\newcommand{\albm}{\alpha}
\newcommand{\alb}{\boldsymbol{\albm}}
\newcommand{\dgnm}{\zeta}
\newcommand{\dgn}{\boldsymbol{\dgnm}}
%synapse models
\newcommand{\wm}{w}
\newcommand{\w}{\mathbf{\wm}}
\newcommand{\Wm}{W}
\newcommand{\W}{\mathbf{\Wm}}
\newcommand{\MMdm}{M}
\newcommand{\MMd}{\mathbf{\MMdm}}
\newcommand{\encm}{K}
\newcommand{\enc}{\mathbf{\encm}}
\newcommand{\frgm}{\Wm^{\mathrm{F}}}
\newcommand{\frg}{\W^{\mathrm{F}}}
\newcommand{\Fm}{\Phi}
\newcommand{\F}{\boldsymbol{\Fm}}
%eigenvectors
\newcommand{\evrm}{u}
\newcommand{\evr}{\mathbf{\evrm}}
\newcommand{\evlm}{\eta}
\newcommand{\evl}{\boldsymbol{\evlm}}
%KKT multipliers
\newcommand{\kktm}{\mu}
\newcommand{\kkt}{\boldsymbol{\kktm}}
%superscripts
\newcommand{\pot}{^{\text{pot}}}
\newcommand{\dep}{^{\text{dep}}}
\newcommand{\potdep}{^{\text{pot/dep}}}
%snr curves etc
\newcommand{\syn}{\vec{w}}
\newcommand{\synid}{\syn_\text{id}}
\DeclareMathOperator{\SNR}{SNR}
\DeclareMathOperator{\snr}{SNR}
\DeclareMathOperator{\NNR}{NNR}
\DeclareMathOperator{\nnr}{NNR}
\newcommand{\snrb}{\overline{\snr}}
\DeclareMathOperator{\env}{Env}
\newcommand{\cnst}{\gamma}
% \newcommand{\CA}{\mathcal{A}}
% \newcommand{\CB}{\mathcal{B}}
% %avoid double superscripts
% \renewcommand{\pdiff}[2]{\frac{\partial #1}{\partial #2}}
% shifted process
\newcommand{\shift}[1]{\widehat{#1}}

@@

We can test if the pattern had been previously stored by computing \(\synid\cdt\syn\) and comparing it to some threshold~\cite{Sommer1998retrieval}.
For large \(N\), this quantity will have a Gausssian distribution.
There will be a ROC curve as a function of this threshold:
%
\begin{equation}\label{eq:ROC}
  \begin{aligned}
  \operatorname{TPR} &= \Phi_c \prn{ \frac{ \Phi_c\inv(\operatorname{FPR}) - \snr }{ \NNR } },
  \quad\text{where } &
    \Phi_c(x) &= \int_x^\infty \frac{ \e^{-\frac{z^2}{2}} }{ \sqrt{2\pi} }\, \dz, \\&&
    \snr &= \frac{ \av{\synid\cdt\syn(t)} - \av{\synid\cdt\syn(\infty)} }{ \sqrt{\var(\synid\cdt\syn(\infty))}}, \\&&
    \NNR &= \sqrt{\frac{ \var(\synid\cdt\syn(t)) }{ \var(\synid\cdt\syn(\infty)) }},
  \end{aligned}
\end{equation}
%
and TPR/FPR are the true/false positive rates.
When the signal-to-noise ratio, \(\snr\) is larger than \(\Phi_c\inv(\operatorname{FPR})\), it is beneficial to decrease the noise-to-noise ratio, NNR.
In the other case, it is beneficial to increase it.
The expectations and variances are over the probability distribution of the synaptic states given the sequence of plasticity events, the probability distribution of the sequence of plasticity events themselves and the probability distribution of the pattern we are testing, \(\synid\).

@@

When the synapse is subjected to a plasticity event the internal state will change stochastically, described by the matrices \(\MMd^\nu\) whose \(i,j\) elements are the transition probabilities from state \(i\) to state \(j\), 
and \(\nu \) labels the diffrerent types of plasticity event.
Potentiating events have \(\nu=\) pot, depressing events have \(\nu=\) dep.
We also use the notation
%
\begin{equation*}
  (-1)^\nu = 
    \begin{cases}
       1, & \nu = \text{pot}, \\
      -1, & \nu = \text{dep}.
    \end{cases}
\end{equation*}
%
The matrices \(\MMd^\nu\) are transition matrices of discrete time Markov processes.
Their elements must satisfy the following constraints
%
\begin{equation}\label{eq:constr}
  \MMdm^\nu_{ij} \geq 0 
    \quad \forall \; \nu, i, j, \qquad
  \sum_j \MMdm^\nu_{ij} = 1
    \quad \forall \; \nu, i.
\end{equation}
%
The upper bounds \(\MMdm^\nu_{ij} \leq 1\) follow automatically from these.

@@

Given a sequence of plasticity events at times \(t_1,t_2,t_3,\ldots\), the probability distribution across the internal states at time \(t_n\), 
mariginalising over states at earlier times, is
%
\begin{equation}\label{eq:plastseq}
  \pr(t_n) = \pr(t_0)  \MMd \pot \MMd \dep \MMd \dep \MMd \pot \MMd \dep \MMd \pot \MMd \pot \MMd \dep \ldots \MMd \pot .
\end{equation}
%
If the probability of a given plasticity event being potentiating or depressing is \(f\pot\) or \(f\dep\) respectively, the probability of the sequence above (conditioned on the number of events) is given by \(f\pot f\dep f\dep f\pot f\dep f\pot f\pot f\dep \ldots f\pot\).

@@

Now to study the SNR as a function of  time, it helps to consider some variables that are more directly related to the time dependance than the matrix elements. 
So we introduce an eigenvector decomposition for \(-\frg\):
%
\begin{equation}\label{eq:eigendecomp}
  \frg = - \sum_a q_a \evr^a \evl^a,
  \quad
  \evl^a \evr^b = \delta_{ab},
  \quad
  \frg \evr^a = -q_a \evr^a,
  \quad
  \evl^a \frg = -q_a \evl^a.
\end{equation}
%
Then we can write
%
\begin{equation}\label{eq:snreig}
  \snr(t) = \sqrt{N}\sum_a \CI_a \e^{-t/\tau_a},
  \qquad
  \snrb(\tau) = \sqrt{N}\sum_a \frac{\CI_a}{1+\tau/\tau_a},
\end{equation}
%
where
%
\begin{equation}\label{eq:snrcoeffs}
  \CI_a = \frac{(\eq \enc \evr^a) (\evl^a \w)}{r},
  \qquad
  \tau_a = \frac{1}{q_a}.
\end{equation}
%
We have ignored the denominator of \cref{eq:SNRcurveExact} for now.
This will not have any effect on the results: it will not change the location of the maxima, which are at \( \Delta f = \eq \w = 0 \).
We will confine such discussion to the footnotes.

@@

The area and \(\Delta p\) are invariant under the scale transformation %\cref{eq:scale}
%
\begin{equation}\label{eq:scale}
  \zeta * \MMd \equiv (1-\zeta) \MMd + \zeta \I .
\end{equation}
%
It follows that
%
\begin{equation*}
  \sum_{\nu, m \neq n} \MMdm^\nu_{mn} \pdiff{\CL}{\MMdm^\nu_{mn}} 
    = - \sum_{\nu, m} \kktm^\nu_{mm} = 0 .
\end{equation*}
%
As the KKT multipliers are nonnegative, this implies that all the diagonal KKT multipliers vanish and we can ignore the diagonal constraints.

@@

We can almost see that degeneracy implies lumpability as follows:
suppose there is a nonzero transition between states \(i\) and \(j\) of type \(\nu\),
Then \(\kktm^\nu_{ij} = 0\) and \cref{eq:areaderivs} becomes
%
\begin{equation}\label{eq:dgn_theta}
\begin{gathered}
  \pdiff{\CL}{\MMdm^\nu_{ij}} =
    r f^\nu \eqm_i \prn{ \brk{\thbm_i - \thbm_j}
     + \prn{c_i + (-1)^\nu + \lambda} \brk{\etwm_i - \etwm_j} } 
    = 0, \\
    \implies \quad
    r f^\nu \eqm_i = 0
    \qquad \text{or} \qquad
    \thbm_i - \thbm_j = -  \prn{c_i + (-1)^\nu + \lambda} \brk{\etwm_i - \etwm_j}. 
\end{gathered}
\end{equation}
%
Therefore, provided that state \(i\) has nonzero equilibrium probability, degeneracy in \(\etw\) implies degeneracy in \(\thb\) for an otimal model.

@@

The fourth inequality can only be saturated if, for each state, the outgoing potentitation rates and depression rates separately sum to one.
This does force degenerate states to be lumpable, and we are left with a serial topology with deterministic transitions.
However, this cannot be done for the two end states without violating the requirements for saturating the first inequality.

@@

Suppose there are two states, \(i,j\), with nonzero transitions from \(m\) for one \(\nu\).
Both must statisfy \cref{eq:oneexit} for the same \(\kktm^\nu_{mm}\), so
%
\begin{equation}\label{eq:twoexits}
  \Wm^\nu_{mi}, \Wm^\nu_{mj} > 0
  \qquad \implies \qquad
  \dgnm^\nu_{m,i} = \dgnm^\nu_{m,j}
  \quad \text{or}\quad
  \dgnm_{m,i} - \dgnm_{m,j} = - (-1)^\nu (\etwm_i - \etwm_j).
\end{equation}
%
Furthermore, for another state \(k\) with zero transition rate from \(m\) for the same \(\nu\):
%
\begin{equation}\label{eq:nonexit}
\begin{aligned}
  \Wm^\nu_{mi} > 0, \Wm^\nu_{mk} = 0 
  \;\;\; \implies \;\;\;
  \kktm^\nu_{mk} = \eqm_m (\dgnm^\nu_{m,k} - \dgnm^\nu_{m,i})
  \;\; \implies \;\;
  \dgnm^\nu_{m,k} &\geq \dgnm^\nu_{m,i} \\
  \text{or}\quad
  \dgnm_{m,k} - \dgnm_{m,i} &\geq - (-1)^\nu (\etwm_k - \etwm_i).
\end{aligned}
\end{equation}
%
If both \(\nu\) have nonzero rates from \(m\) to \(i\) and \(j\):
%
\begin{equation}\label{eq:twodoubleexits}
  \Wm\pot_{mi}, \Wm\dep_{mi}, \Wm\pot_{mj}, \Wm\dep_{mj} > 0
  \qquad \implies \qquad
  \dgnm_{m,i} = \dgnm_{m,j} 
  \quad \text{and} \quad
  \etwm_i = \etwm_j.
\end{equation}
%
The consequences of these inequalities are depicted in \cref{fig:planes}(\ref{fig:triangle_spacing}).
We see that \(n\) must be a global minimum of \(\dgn_m\), and if there are multiple transitions of a single type (potentiating or depressing) the targets are constrained to lie on a single line.
Even if there is no state \(n\) that is the target of both potentiating and depressing transitions, there will be two lines at \(45^\circ \) passing through any targets of potentiating/depressing transitions that demarcate the forbidden zone.

We use \(m*\) to label the intersection of the two diagonal lines in \cref{fig:planes}(\ref{fig:triangle_spacing}), wether or not there is a state there.
It can be defined for each \(m\), provided that it has at least one outgoing transition for each plasticity type, via
%
\begin{equation*}
\begin{lgathered}  
  \dgnm\pot_{m*} = \dgnm\pot_{m,j} = \min_{k\neq m} \dgnm\pot_{m,k},
  \qquad
  \dgnm\dep_{m*} = \dgnm\dep_{m,i} = \min_{k\neq m} \dgnm\dep_{m,k},
  \\
  \text{and also} \quad
  \dgnm_{m*} = \frac{1}{2} \prn{\dgnm\pot_{m*} + \dgnm\dep_{m*}},
  \quad
  \etwm_{m*} = \frac{1}{2} \prn{\dgnm\pot_{m*} - \dgnm\dep_{m*}},
  \quad
  \thbm_{m*} = \dgnm_{m*} - c_m \etwm_{m*}.
\end{lgathered}
\end{equation*}
%
As shown in \cref{eq:twoexits}, this produces the same point regardless of which states \(i\) and \(j\) we use, 
so long as there is a nonzero depressing/potentiating transition from \(m\) to that state.
The KKT multipliers are related to these quantities by
%
\begin{equation}\label{eq:KKTz}
  \kktm^\nu_{mm} = \eqm_m (\dgnm^\nu_{m,m} - \dgnm^\nu_{m*}),
  \qquad
  \kktm^\nu_{mk} = \eqm_m (\dgnm^\nu_{m,k} - \dgnm^\nu_{m*}).
\end{equation}
%

@@

From \cref{sec:serialarea}, we know the optimal models at \(s = 0\).
% (There is still an ambiguity concering the scale, but it makes sense to choose to saturate the diagonal constraints in \cref{eq:wdof} so that \cref{eq:optmodelscale} holds).
At \(r = 0\) there is only one allowed model: \(\W = \mathbf{0}\).

@@

We can maximise this \wrt \(\varepsilon\), which yields
%
\begin{equation}\label{eq:shortopt}
  (\varepsilon - f_m)^2 = (m - f_m)(g_m - f_m),
  \quad \implies \quad
  \hat{A}(s) = \frac{\Delta f_{m}}
      {s \Delta g_{m} \brk{\sqrt{\abs{m - f_m}} - \sqrt{\abs{g_m - f_m}}}^2}.
\end{equation}
%
We can find the boundary of the region of applicability by asking at what value of \(s\) is the optimal value of \(\varepsilon=0\) or \(1\)?
%
\begin{equation}\label{eq:shortenvalid}
\begin{aligned}
  \left.\pdiff{\hat{A}(s)}{\varepsilon}\right|_{\mathrlap{\varepsilon=0}} \propto &\,
    m \zeta \brk{S(m\beta) - S((m-1)\beta)} - S(m\beta) [S(m\beta) + 1]
      = 0,\\
  \left.\pdiff{\hat{A}(s)}{\varepsilon}\right|_{\mathrlap{\varepsilon=1}} \propto &\,
    (m-1)\brk{S(m\beta) - S((m-1)\beta)} - \zeta S((m-1)\beta) [S((m-1)\beta) + 1]
      = 0,\\
    &\text{where}\quad
    \zeta = 2s + 1, \qquad
    s = S(\beta) = 2 \sinh^2(\beta/2),
\end{aligned}
\end{equation}
%
with the \(\varepsilon=0\) case providing the lower boundary and \(\varepsilon=1\) providing the upper boundary.
For lower \(s\), \(\varepsilon = 0\) is the optimum. 
For higher \(s\), \(\varepsilon = 1\) is the optimum, which is eqivalent to \(\varepsilon = 0\) with \(M-2\) states.
This is depicted by the dotted lines in \cref{fig:shorten}\ref{fig:short_ratio}.

>>>>

\newcommand{\aav}[1]{\mathopen{}\left\llangle #1 \right\rrangle\mathclose{}}
\newcommand{\nnrm}[1]{\mathopen{}\left\lWavy #1 \right\rWavy\mathclose{}}
%================================================================================
%
% \usepackage{adjustbox}
% \newcommand{\aligntop}[1]{\adjustbox{valign=t}{#1}}
% \newcommand{\alignmid}[1]{\adjustbox{valign=c}{#1}}
% %
% \usepackage{paralist}
% \newenvironment{myenuma}{\begin{inparaenum}[(a)]}{\end{inparaenum}}
% \newenvironment{myenumi}{\begin{inparaenum}[i)]}{\end{inparaenum}}
% \newenvironment{myenumA}{\begin{inparaenum}[\bfseries A.]}{\end{inparaenum}}
% \newenvironment{myenumI}{\begin{inparaenum}[\bfseries I.]}{\end{inparaenum}}
% \newenvironment{myenumr}{\begin{inparaenum}[1.]}{\end{inparaenum}}
% %for nested lists, \ref only produces innermost item:  (hacky)
% \makeatletter % for internal macros with @
% \renewcommand\p@enumii{}
% \makeatother
%
%================================================================================
%matrices
%vec of ones
\newcommand{\onev}{\mathbf{e}}
%mat of ones
\newcommand{\onem}{\mathbf{E}}
%Markov matrix
\newcommand{\MMm}{Q}
\newcommand{\MM}{\mathbf{\MMm}}
%probability distribution
\newcommand{\prm}{p}
\newcommand{\pr}{\mathbf{\prm}}
%equilibrium distribution
\newcommand{\eqm}{\pi}
\newcommand{\eq}{\boldsymbol{\eqm}}
%equilibrium metric
\newcommand{\Eqm}{\Pi}
\newcommand{\Eq}{\boldsymbol{\Eqm}}
%first passage times
\newcommand{\fptm}{T}
\newcommand{\fpt}{\mathbf{\fptm}}
%off-diagonal first passage times
\newcommand{\fptbm}{\overline{\fptm}}
\newcommand{\fptb}{\overline{\fpt}}
%fundamental matrix
\newcommand{\fundm}{Z}
\newcommand{\fund}{\mathbf{\fundm}}
%other symbols
\newcommand{\Pb}{\mathbf{P}}
\newcommand{\pib}{\boldsymbol{\pi}}
\newcommand{\arowm}{\xi}
\newcommand{\arow}{\boldsymbol{\arowm}}
\newcommand{\Lbm}{\Lambda}
\newcommand{\Lb}{\boldsymbol{\Lbm}}
\newcommand{\lbm}{\lambda}
\newcommand{\lb}{\boldsymbol{\lbm}}
\newcommand{\etwm}{\eta^w}
\newcommand{\etw}{\boldsymbol{\eta}^w}
\newcommand{\thbm}{\theta}
\newcommand{\thb}{\boldsymbol{\thbm}}
\newcommand{\ombm}{\omega}
\newcommand{\omb}{\boldsymbol{\ombm}}
\newcommand{\albm}{\alpha}
\newcommand{\alb}{\boldsymbol{\albm}}
\newcommand{\dgnm}{\zeta}
\newcommand{\dgn}{\boldsymbol{\dgnm}}
%synapse models
\newcommand{\wm}{w}
\newcommand{\w}{\mathbf{\wm}}
\newcommand{\Wm}{W}
\newcommand{\W}{\mathbf{\Wm}}
\newcommand{\MMdm}{M}
\newcommand{\MMd}{\mathbf{\MMdm}}
\newcommand{\encm}{K}
\newcommand{\enc}{\mathbf{\encm}}
\newcommand{\frgm}{\Wm^{\mathrm{F}}}
\newcommand{\frg}{\W^{\mathrm{F}}}
\newcommand{\Fm}{\Phi}
\newcommand{\F}{\boldsymbol{\Fm}}
%eigenvectors
\newcommand{\evrm}{u}
\newcommand{\evr}{\mathbf{\evrm}}
\newcommand{\evlm}{\eta}
\newcommand{\evl}{\boldsymbol{\evlm}}
%KKT multipliers
\newcommand{\kktm}{\mu}
\newcommand{\kkt}{\boldsymbol{\kktm}}
%superscripts
\newcommand{\pot}{^{\text{pot}}}
\newcommand{\dep}{^{\text{dep}}}
\newcommand{\potdep}{^{\text{pot/dep}}}
%snr curves etc
\newcommand{\syn}{\vec{w}}
\newcommand{\synid}{\syn_\text{id}}
\DeclareMathOperator{\SNR}{SNR}
\DeclareMathOperator{\snr}{SNR}
\DeclareMathOperator{\NNR}{NNR}
\DeclareMathOperator{\nnr}{NNR}
\newcommand{\snrb}{\overline{\snr}}
\DeclareMathOperator{\env}{Env}
\newcommand{\cnst}{\gamma}
% \newcommand{\CA}{\mathcal{A}}
% \newcommand{\CB}{\mathcal{B}}
% %avoid double superscripts
% \renewcommand{\pdiff}[2]{\frac{\partial #1}{\partial #2}}
% shifted process
\newcommand{\shift}[1]{\widehat{#1}}

@@

We can test if the pattern had been previously stored by computing \(\synid\cdt\syn\) and comparing it to some threshold~\cite{Sommer1998retrieval}.
For large \(N\), this quantity will have a Gaussian distribution.
There will be a ROC curve as a function of this threshold:
%
\begin{equation}\label{eq:ROC}
  \begin{aligned}
  \operatorname{TPR} &= \Phi_c \prn{ \frac{ \Phi_c\inv(\operatorname{FPR}) - \snr }{ \NNR } },
  \quad\text{where } &
    \Phi_c(x) &= \int_x^\infty \frac{ \e^{-\frac{z^2}{2}} }{ \sqrt{2\pi} }\, \dz, \\&&
    \snr &= \frac{ \av{\synid\cdt\syn(t)} - \av{\synid\cdt\syn(\infty)} }{ \sqrt{\var(\synid\cdt\syn(\infty))}}, \\&&
    \NNR &= \sqrt{\frac{ \var(\synid\cdt\syn(t)) }{ \var(\synid\cdt\syn(\infty)) }},
  \end{aligned}
\end{equation}
%
and TPR/FPR are the true/false positive rates.
When the signal-to-noise ratio, \(\snr\) is larger than \(\Phi_c\inv(\operatorname{FPR})\), it is beneficial to decrease the noise-to-noise ratio, NNR.
In the other case, it is beneficial to increase it.
The expectations and variances are over the probability distribution of the synaptic states given the sequence of plasticity events, the probability distribution of the sequence of plasticity events themselves and the probability distribution of the pattern we are testing, \(\synid\).

@@

When the synapse is subjected to a plasticity event the internal state will change stochastically, described by the matrices \(\MMd^\nu\) whose \(i,j\) elements are the transition probabilities from state \(i\) to state \(j\), 
and \(\nu \) labels the different types of plasticity event.
Potentiating events have \(\nu=\) pot, depressing events have \(\nu=\) dep.
We also use the notation
%
\begin{equation*}
  (-1)^\nu = 
    \begin{cases}
       1, & \nu = \text{pot}, \\
      -1, & \nu = \text{dep}.
    \end{cases}
\end{equation*}
%
The matrices \(\MMd^\nu\) are transition matrices of discrete time Markov processes.
Their elements must satisfy the following constraints
%
\begin{equation}\label{eq:constr}
  \MMdm^\nu_{ij} \geq 0 
    \quad \forall \; \nu, i, j, \qquad
  \sum_j \MMdm^\nu_{ij} = 1
    \quad \forall \; \nu, i.
\end{equation}
%
The upper bounds \(\MMdm^\nu_{ij} \leq 1\) follow automatically from these.

@@

Given a sequence of plasticity events at times \(t_1,t_2,t_3,\ldots\), the probability distribution across the internal states at time \(t_n\), 
marginalising over states at earlier times, is
%
\begin{equation}\label{eq:plastseq}
  \pr(t_n) = \pr(t_0)  \MMd \pot \MMd \dep \MMd \dep \MMd \pot \MMd \dep \MMd \pot \MMd \pot \MMd \dep \ldots \MMd \pot .
\end{equation}
%
If the probability of a given plasticity event being potentiating or depressing is \(f\pot\) or \(f\dep\) respectively, the probability of the sequence above (conditioned on the number of events) is given by \(f\pot f\dep f\dep f\pot f\dep f\pot f\pot f\dep \ldots f\pot\).

@@

Now to study the SNR as a function of  time, it helps to consider some variables that are more directly related to the time dependence than the matrix elements. 
So we introduce an eigenvector decomposition for \(-\frg\):
%
\begin{equation}\label{eq:eigendecomp}
  \frg = - \sum_a q_a \evr^a \evl^a,
  \quad
  \evl^a \evr^b = \delta_{ab},
  \quad
  \frg \evr^a = -q_a \evr^a,
  \quad
  \evl^a \frg = -q_a \evl^a.
\end{equation}
%
Then we can write
%
\begin{equation}\label{eq:snreig}
  \snr(t) = \sqrt{N}\sum_a \CI_a \e^{-t/\tau_a},
  \qquad
  \snrb(\tau) = \sqrt{N}\sum_a \frac{\CI_a}{1+\tau/\tau_a},
\end{equation}
%
where
%
\begin{equation}\label{eq:snrcoeffs}
  \CI_a = \frac{(\eq \enc \evr^a) (\evl^a \w)}{r},
  \qquad
  \tau_a = \frac{1}{q_a}.
\end{equation}
%
We have ignored the denominator of \cref{eq:SNRcurveExact} for now.
This will not have any effect on the results: it will not change the location of the maxima, which are at \( \Delta f = \eq \w = 0 \).
We will confine such discussion to the footnotes.

@@

The area and \(\Delta p\) are invariant under the scale transformation %\cref{eq:scale}
%
\begin{equation}\label{eq:scale}
  \zeta * \MMd \equiv (1-\zeta) \MMd + \zeta \I .
\end{equation}
%
It follows that
%
\begin{equation*}
  \sum_{\nu, m \neq n} \MMdm^\nu_{mn} \pdiff{\CL}{\MMdm^\nu_{mn}} 
    = - \sum_{\nu, m} \kktm^\nu_{mm} = 0 .
\end{equation*}
%
As the KKT multipliers are non-negative, this implies that all the diagonal KKT multipliers vanish and we can ignore the diagonal constraints.

@@

We can almost see that degeneracy implies lumpability as follows:
suppose there is a nonzero transition between states \(i\) and \(j\) of type \(\nu\),
Then \(\kktm^\nu_{ij} = 0\) and \cref{eq:areaderivs} becomes
%
\begin{equation}\label{eq:dgn_theta}
\begin{gathered}
  \pdiff{\CL}{\MMdm^\nu_{ij}} =
    r f^\nu \eqm_i \prn{ \brk{\thbm_i - \thbm_j}
     + \prn{c_i + (-1)^\nu + \lambda} \brk{\etwm_i - \etwm_j} } 
    = 0, \\
    \implies \quad
    r f^\nu \eqm_i = 0
    \qquad \text{or} \qquad
    \thbm_i - \thbm_j = -  \prn{c_i + (-1)^\nu + \lambda} \brk{\etwm_i - \etwm_j}. 
\end{gathered}
\end{equation}
%
Therefore, provided that state \(i\) has nonzero equilibrium probability, degeneracy in \(\etw\) implies degeneracy in \(\thb\) for an optimal model.

@@

The fourth inequality can only be saturated if, for each state, the outgoing potentiation rates and depression rates separately sum to one.
This does force degenerate states to be lumpable, and we are left with a serial topology with deterministic transitions.
However, this cannot be done for the two end states without violating the requirements for saturating the first inequality.

@@

Suppose there are two states, \(i,j\), with nonzero transitions from \(m\) for one \(\nu\).
Both must satisfy \cref{eq:oneexit} for the same \(\kktm^\nu_{mm}\), so
%
\begin{equation}\label{eq:twoexits}
  \Wm^\nu_{mi}, \Wm^\nu_{mj} > 0
  \qquad \implies \qquad
  \dgnm^\nu_{m,i} = \dgnm^\nu_{m,j}
  \quad \text{or}\quad
  \dgnm_{m,i} - \dgnm_{m,j} = - (-1)^\nu (\etwm_i - \etwm_j).
\end{equation}
%
Furthermore, for another state \(k\) with zero transition rate from \(m\) for the same \(\nu\):
%
\begin{equation}\label{eq:nonexit}
\begin{aligned}
  \Wm^\nu_{mi} > 0, \Wm^\nu_{mk} = 0 
  \;\;\; \implies \;\;\;
  \kktm^\nu_{mk} = \eqm_m (\dgnm^\nu_{m,k} - \dgnm^\nu_{m,i})
  \;\; \implies \;\;
  \dgnm^\nu_{m,k} &\geq \dgnm^\nu_{m,i} \\
  \text{or}\quad
  \dgnm_{m,k} - \dgnm_{m,i} &\geq - (-1)^\nu (\etwm_k - \etwm_i).
\end{aligned}
\end{equation}
%
If both \(\nu\) have nonzero rates from \(m\) to \(i\) and \(j\):
%
\begin{equation}\label{eq:twodoubleexits}
  \Wm\pot_{mi}, \Wm\dep_{mi}, \Wm\pot_{mj}, \Wm\dep_{mj} > 0
  \qquad \implies \qquad
  \dgnm_{m,i} = \dgnm_{m,j} 
  \quad \text{and} \quad
  \etwm_i = \etwm_j.
\end{equation}
%
The consequences of these inequalities are depicted in \cref{fig:planes}(\ref{fig:triangle_spacing}).
We see that \(n\) must be a global minimum of \(\dgn_m\), and if there are multiple transitions of a single type (potentiating or depressing) the targets are constrained to lie on a single line.
Even if there is no state \(n\) that is the target of both potentiating and depressing transitions, there will be two lines at \(45^\circ \) passing through any targets of potentiating/depressing transitions that demarcate the forbidden zone.

We use \(m*\) to label the intersection of the two diagonal lines in \cref{fig:planes}(\ref{fig:triangle_spacing}), whether or not there is a state there.
It can be defined for each \(m\), provided that it has at least one outgoing transition for each plasticity type, via
%
\begin{equation*}
\begin{lgathered}  
  \dgnm\pot_{m*} = \dgnm\pot_{m,j} = \min_{k\neq m} \dgnm\pot_{m,k},
  \qquad
  \dgnm\dep_{m*} = \dgnm\dep_{m,i} = \min_{k\neq m} \dgnm\dep_{m,k},
  \\
  \text{and also} \quad
  \dgnm_{m*} = \frac{1}{2} \prn{\dgnm\pot_{m*} + \dgnm\dep_{m*}},
  \quad
  \etwm_{m*} = \frac{1}{2} \prn{\dgnm\pot_{m*} - \dgnm\dep_{m*}},
  \quad
  \thbm_{m*} = \dgnm_{m*} - c_m \etwm_{m*}.
\end{lgathered}
\end{equation*}
%
As shown in \cref{eq:twoexits}, this produces the same point regardless of which states \(i\) and \(j\) we use, 
so long as there is a nonzero depressing/potentiating transition from \(m\) to that state.
The KKT multipliers are related to these quantities by
%
\begin{equation}\label{eq:KKTz}
  \kktm^\nu_{mm} = \eqm_m (\dgnm^\nu_{m,m} - \dgnm^\nu_{m*}),
  \qquad
  \kktm^\nu_{mk} = \eqm_m (\dgnm^\nu_{m,k} - \dgnm^\nu_{m*}).
\end{equation}
%

@@

From \cref{sec:serialarea}, we know the optimal models at \(s = 0\).
% (There is still an ambiguity concerning the scale, but it makes sense to choose to saturate the diagonal constraints in \cref{eq:wdof} so that \cref{eq:optmodelscale} holds).
At \(r = 0\) there is only one allowed model: \(\W = \mathbf{0}\).

@@

We can maximise this \wrt \(\varepsilon\), which yields
%
\begin{equation}\label{eq:shortopt}
  (\varepsilon - f_m)^2 = (m - f_m)(g_m - f_m),
  \quad \implies \quad
  \hat{A}(s) = \frac{\Delta f_{m}}
      {s \Delta g_{m} \brk{\sqrt{\abs{m - f_m}} - \sqrt{\abs{g_m - f_m}}}^2}.
\end{equation}
%
We can find the boundary of the region of applicability by asking at what value of \(s\) is the optimal value of \(\varepsilon=0\) or \(1\)?
%
\begin{equation}\label{eq:shortenvalid}
\begin{aligned}
  \left.\pdiff{\hat{A}(s)}{\varepsilon}\right|_{\mathrlap{\varepsilon=0}} \propto &\,
    m \zeta \brk{S(m\beta) - S((m-1)\beta)} - S(m\beta) [S(m\beta) + 1]
      = 0,\\
  \left.\pdiff{\hat{A}(s)}{\varepsilon}\right|_{\mathrlap{\varepsilon=1}} \propto &\,
    (m-1)\brk{S(m\beta) - S((m-1)\beta)} - \zeta S((m-1)\beta) [S((m-1)\beta) + 1]
      = 0,\\
    &\text{where}\quad
    \zeta = 2s + 1, \qquad
    s = S(\beta) = 2 \sinh^2(\beta/2),
\end{aligned}
\end{equation}
%
with the \(\varepsilon=0\) case providing the lower boundary and \(\varepsilon=1\) providing the upper boundary.
For lower \(s\), \(\varepsilon = 0\) is the optimum. 
For higher \(s\), \(\varepsilon = 1\) is the optimum, which is equivalent to \(\varepsilon = 0\) with \(M-2\) states.
This is depicted by the dotted lines in \cref{fig:shorten}\ref{fig:short_ratio}.
